---
title: "Exascalar: Visualizing change in the Top500 and Green500"
author: "Winston"
date: "November 27, 2014"
output: html_document
---

###Comparing the November and June 2014 Green500 and Top500

Exascalar is one of the best ways to visualize changes in the [Green500](http:\\green500.org) and [Top500](http:\\top500.org) lists.  
While the #1 supercomputer in the Top500 has not changed in the last four publications of the lists, the population of the Top500 list has been far from stagnant. For Instance, the Green500 list has been quite dynamic. 


```{r, echo=FALSE}

## This program imports cleaned data from the Green500 and Top500 lists

## there are several subtleties of the implenetation here which need to be paid attention to when adapting it.
##each "500" list is called out explicitly
##in a couple of places the number of lists (currently 10) is assumed.

## GET THE CLEANED DATA

##check for Exascalar Directory. If none exists stop program with error

##check to ensure results director exists

if(getwd()!="/Users/winstonsaunders/Documents/Exascalar") setwd("/Users/winstonsaunders/Documents/Exascalar")

#if(!file.exists("./results")) stop("Data not found in directory Exascalar, first run Exascalar_Cleaner to get tidy data")


## set working directory


# define Data Directories to use
results <- "./results"

## ------------------------
## Read results files

# import data set

## there are probably ways to simplify this code but this brute force method is easy to read.

Nov14 <- read.csv(paste0(results, "/Nov14.csv"), header=TRUE)
Jun14 <- read.csv(paste0(results, "/Jun14.csv"), header=TRUE)
Nov13 <- read.csv(paste0(results, "/Nov13.csv"), header=TRUE)
Jun13 <- read.csv(paste0(results, "/Jun13.csv"), header=TRUE)
Nov12 <- read.csv(paste0(results, "/Nov12.csv"), header=TRUE)
Jun12 <- read.csv(paste0(results, "/Jun12.csv"), header=TRUE)
Nov11 <- read.csv(paste0(results, "/Nov11.csv"), header=TRUE)
Jun11 <- read.csv(paste0(results, "/Jun11.csv"), header=TRUE)
Nov10 <- read.csv(paste0(results, "/Nov10.csv"), header=TRUE)
Jun10 <- read.csv(paste0(results, "/Jun10.csv"), header=TRUE)
Nov09 <- read.csv(paste0(results, "/Nov09.csv"), header=TRUE)
Jun09 <- read.csv(paste0(results, "/Jun09.csv"), header=TRUE)
#Nov08 <- read.csv(paste0(results, "/green500_top_200811.csv"), header=TRUE)
#Jun08 <- read.csv(paste0(results, "/green500_top_200806.csv"), header=TRUE)


#print(Jun14[1:10, "ExaRank", "top500rank", "green500rank", computername], row.names=FALSE)

```


### Exascalar Plot 

The easiest way to visualize change in the Top500 and Green500 lists is to just overlay plot from November 2014 with that of June 2014.  

In the plot below points from November are smaller red dots, points with empty blue circles are computers that are no longer on the list, and red points with blue circles around them are computers on both lists.  

```{r, echo=FALSE, fig.align='center', fig.height=6, fig.width=5}
## EXASCALAR PLOT OVERLAYING TWO LISTS

## plots "reference" list first, then "list of current interest" is overlayed

plot(Jun14$mflopswatt ,
     Jun14$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "Exascalar Delta", 
     col = "blue",
     cex=1,
     pch=1, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))


par(new=TRUE)

plot(Nov14$mflopswatt ,
     Nov14$rmax*10^3, 
     log="xy", 
     asp = TRUE, 
     xlab = "",
     ylab = "", 
     main = "", 
     col = "red",
     bg = "steelblue2",
     cex=1,
     pch=20, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))





##add text to plots  (Some are commented out to clean up appearance, but left in for possible later convenience)

legend(4E1,1E11, legend = c("Nov 14","Jun 14"), pch=c(20,1), col = c("red", "blue"), 
       pt.bg=c("steelblue", NULL), cex=0.8)

text(1.0E4,
     5.E7, "Nov 2014", cex=.7, col="black", pos=3)
text(0.8E4,
     3.3E7, "data from Green500 and Top500 Lists", cex=.4, col="black", pos=3)


#text(0.35e+05, 2e+12, expression(epsilon == 0), cex=.7, srt=-45)
text(0.35e+04, 1e+12, expression(epsilon == -1), cex=.7, srt=-45)
text(1.2e+02, 1e+12, expression(epsilon == -2), cex=.7, srt=-45)
text(0.2e+02, .25e+12, expression(epsilon == -3), cex=.7, srt=-45)
text(0.15e+02, 1.3e+10, expression(epsilon == -4), cex=.7, srt=-45)

text(0.15e+05, .5e+12, "20 MWatt", cex=.7, srt=45)
text(0.15e+05, .5e+10, "0.2 MWatt", cex=.7, srt=45)
#text(0.15e+05, .5e+08, "0.02 MWatt", cex=.7, srt=45)


## GENERATE CONSTANT POWER AND ISO-EXASCALAR LINES FOR THE GRAPH

## These are tbe hashed lines on the graph representing "constant power" and "constant exascalar"
## The approach here is to define functions that create two endpoints for the line segments that 
##    depend on power and exascalar, respectively

isopowerline <- function(megawatts, efficiencyrangelow=10*.9, efficiencyrangehigh=100000*1.1, perfrangehigh = 2*10^12*1.1, perfrangelow = 0.9**5*10^6)
{matrix(c(efficiencyrangelow, efficiencyrangehigh, efficiencyrangelow*megawatts*10^6, efficiencyrangehigh*megawatts*10^6), ncol=2)
}

## Compute Graphical Lines for iso power
## the argument passed in isopowerline is in megawatts

for (i in -2:4) {
        lines(isopowerline(2*10^-i)[,1],isopowerline(2*10^-i)[,2], lwd=.5, lty=2)}

## add graphical lines for iso "exscalar"


ExaPerf <- 10^12           ##in Megaflops
ExaEff <- 10^12/(20*10^6)  ##in Megaflops/Watt

Exascale = ExaPerf*ExaEff

isoexaline <- function(exascalar, efficiencyrangelow=0.5*20, efficiencyrangehigh=2*50000, perfrangehigh = 2*10^12, perfrangelow = 0.5*5*10^7){
        matrix(c(Exascale/(10^(-exascalar*sqrt(2)))/perfrangehigh, efficiencyrangehigh, perfrangehigh, Exascale/(10^(-exascalar*sqrt(2)))/efficiencyrangehigh), ncol=2)
}

## Compute Graphical Lines for iso exascalar
## the argument passed is - log10 of exascalar

for (i in 0:7) {lines(isoexaline(-i)[,1], isoexaline(-i)[,2], lwd=.5, lty=2)}



```

####Overall List

The __Top Performance__ supercomputer, 
`r Nov14[Nov14$rmax==max(Nov14$rmax), "computer"]` 
with a performance of `r Nov14[Nov14$rmax==max(Nov14$rmax), "rmax"]` mflops
and efficiency of `r Nov14[Nov14$rmax==max(Nov14$rmax), "mflopswatt"]` mflops/watt.   
Exascalar rank `r Nov14[Nov14$rmax==max(Nov14$rmax), "ExaRank"]`    
Top500 rank of `r Nov14[Nov14$rmax==max(Nov14$rmax), "top500rank"]`   
Green500 rank of `r Nov14[Nov14$rmax==max(Nov14$rmax), "green500rank"]`   

The __Top Efficiency__ supercomputer, 
`r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "computer"]`  
with a performance of `r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "rmax"]` mflops 
and efficiency of `r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "mflopswatt"]`     
Exascalar rank of `r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "ExaRank"]`   
Top500 rank of `r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "top500rank"]`    
Green500 rank of `r Nov14[Nov14$mflopswatt==max(Nov14$mflopswatt), "green500rank"]`  

The __Lowest Power__ supercomputer, 
`r Nov14[Nov14$power==min(Nov14$power), "computer"]`, 
with a performance of `r Nov14[Nov14$power==min(Nov14$power), "rmax"]` mflops
and efficiency of `r Nov14[Nov14$power==min(Nov14$power), "mflopswatt"]` mflops/watt 
and power of  `r Nov14[Nov14$power==min(Nov14$power), "power"]` kWatt    
Top500 rank of `r Nov14[Nov14$power==min(Nov14$power), "top500rank"]`


The __Highest Power__ supercomputer `r Nov14[Nov14$power==max(Nov14$power), "computer"]` 
consumes `r Nov14[Nov14$power==max(Nov14$power), "power"]` kWatt, has an Exascalar rank of `r Nov14[Nov14$power==max(Nov14$power), "ExaRank"]`    
Top500rank of `r Nov14[Nov14$power==max(Nov14$power), "top500rank"]`.  




####Interesting new entrants



```{r, echo=FALSE} 





NewList<-Nov14[!(Nov14$exascalar %in% Jun14$exascalar),]





```

Of the _new_ computers on the list there are some interesting entries that I capture below:

The __Top Performing__ NEW Supercomputer is  
`r NewList[NewList$rmax==max(NewList$rmax), "computer"]`  
with a performance of `r NewList[NewList$rmax==max(NewList$rmax), "rmax"]` mflops  
an efficiency of `r NewList[NewList$rmax==max(NewList$rmax), "mflopswatt"]` mflops/watt  
Exascalar Rank of `r NewList[NewList$rmax==max(NewList$rmax), "ExaRank"]`  
Top500 rank of `r NewList[NewList$rmax==max(NewList$rmax), "top500rank"]`  
Green500 rank of `r NewList[NewList$rmax==max(NewList$rmax), "green500rank"]`

The __Top Efficiency__ NEW Supercomputer is  
`r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "computer"]`  
with a performance of `r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "rmax"]` mflops and an efficiency of `r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "mflopswatt"]` mflops/watt.  
Exascalar Rank of `r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "ExaRank"]`  
Top500 rank of `r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "top500rank"]`  
Green500 rank of `r NewList[NewList$mflopswatt==max(NewList$mflopswatt), "green500rank"]`

The __Lowest Power__ NEW Supercomputer is  
`r NewList[NewList$power==min(NewList$power), "computer"]`  
with a performance of `r NewList[NewList$power==min(NewList$power), "rmax"]` mflops
and efficiency of `r NewList[NewList$power==min(NewList$power), "mflopswatt"]` mflops/watt.  
Its power is `r NewList[NewList$power==min(NewList$power), "power"]` kWatt.   
Exascalar rank is `r NewList[NewList$power==min(NewList$power), "ExaRank"]`  
Top500 rank is `r NewList[NewList$power==min(NewList$power), "top500rank"]`  
Green500rank `r NewList[NewList$power==min(NewList$power), "green500rank"]` 


The __Highest Power__ NEW Supercomputer   
`r NewList[NewList$power==max(NewList$power), "computer"]`  
Consumes `r NewList[NewList$power==max(NewList$power), "power"]` kWatt 
with a performance of `r NewList[NewList$power==max(NewList$power), "rmax"]` mflops and has an Exascalar Rank of  `r NewList[NewList$power==max(NewList$power), "ExaRank"]`  

####Highest and Lowest Ranked New Computers


```{r, echo=FALSE}

###Shorten Name to fit on page

NewList$minrank<- apply(NewList[,c(3,5,6)],1,min)
NewList$maxrank<- apply(NewList[,c(3,5,6)],1,max)


TopNew<-NewList[with(NewList, order(minrank)), ]
TopNew$computer<-substring(TopNew$computer,1,40)


BottomNew<-NewList[with(NewList, order(-maxrank)), ]
BottomNew$computer<-substring(BottomNew$computer,1,40)


```

The highest ranked new computers on the list are  
```{r, echo=FALSE}
print(TopNew[1:10, c(3,5,6,10)], row.names=FALSE)
```

The median Exascalar of the New Computers is `r round(median(NewList$exascalar),2)` compared to the median of all computers of `r round(median(Nov14$exascalar),2)`.

Surprisingly there are new entrants on the list that have relatively low rankings, among these 
```{r, echo=FALSE}
print(BottomNew[1:10, c(3,5,6,10)], row.names=FALSE)
```



.



