---
title: "Visualizing Top500/Green500 Influencers with Exascalar"
author: "Winston Saunders"
date: "December 2, 2014"
output: 
  html_document:
    keep_md: true
---

###Visulizing Key Trends

Recall Exascalar is one of the best ways to visualize changes in the [Green500](http:\\green500.org) and [Top500](http:\\top500.org) lists since it compares both system performance and efficiency in one common analysis. To the extent that performance and efficiency trends are ultimately related, Exascalar enables an overview of population trends.

In this analysis I explore using Exascalar as a visualization tool to look at architectural influencers efficiency and performance. This analysis was suggested in a tweet from the folks at [HPC Wire](http://www.hpcwire.com/).

Please note: This is a first cut at this analysis. In the future I plan to migrate plotting to ggplot2 and add some interactive features. If you have comments please tweet them to me @WinstonOnEnergy.


```{r, echo=FALSE}

## This program imports cleaned data from the Green500 and Top500 lists

## there are several subtleties of the implenetation here which need to be paid attention to when adapting it.
##each "500" list is called out explicitly
##in a couple of places the number of lists (currently 10) is assumed.

## GET THE CLEANED DATA

##check for Exascalar Directory. If none exists stop program with error

##check to ensure results director exists

if(getwd()!="/Users/winstonsaunders/Documents/Exascalar") setwd("/Users/winstonsaunders/Documents/Exascalar")

#if(!file.exists("./results")) stop("Data not found in directory Exascalar, first run Exascalar_Cleaner to get tidy data")


## set working directory


# define Data Directories to use
results <- "./results"

## ------------------------
## Read results files

# import data set

## there are probably ways to simplify this code but this brute force method is easy to read.

Nov14 <- read.csv(paste0(results, "/Nov14.csv"), header=TRUE)
Jun14 <- read.csv(paste0(results, "/Jun14.csv"), header=TRUE)
Nov13 <- read.csv(paste0(results, "/Nov13.csv"), header=TRUE)
Jun13 <- read.csv(paste0(results, "/Jun13.csv"), header=TRUE)
Nov12 <- read.csv(paste0(results, "/Nov12.csv"), header=TRUE)
Jun12 <- read.csv(paste0(results, "/Jun12.csv"), header=TRUE)
Nov11 <- read.csv(paste0(results, "/Nov11.csv"), header=TRUE)
Jun11 <- read.csv(paste0(results, "/Jun11.csv"), header=TRUE)
Nov10 <- read.csv(paste0(results, "/Nov10.csv"), header=TRUE)
Jun10 <- read.csv(paste0(results, "/Jun10.csv"), header=TRUE)
Nov09 <- read.csv(paste0(results, "/Nov09.csv"), header=TRUE)
Jun09 <- read.csv(paste0(results, "/Jun09.csv"), header=TRUE)


```


### November 2014 Exascalar Plot 

For reference here is the November 2014 Exascalar Plot.


  
```{r, echo=FALSE, fig.align='center', fig.height=7.25, fig.width=6}
## EXASCALAR PLOT OVERLAYING TWO LISTS

## plots "reference" list first, then "list of current interest" is overlayed

plot(Nov14$mflopswatt ,
     Nov14$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2014 Exascalar", 
     col = "red",
     cex=1.1,
     pch=20, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))



##add text to plots  (Some are commented out to clean up appearance, but left in for possible later convenience)

legend(4E1,1E11, legend = "Nov 14", pch=20, col = "red", cex=0.8)

text(1.0E4,
     5.E7, "Nov 2014", cex=.7, col="black", pos=3)
text(0.8E4,
     3.3E7, "data from Green500 and Top500 Lists", cex=.7, col="black", pos=3)


#text(0.35e+05, 2e+12, expression(epsilon == 0), cex=.7, srt=-45)
text(0.35e+04, 1e+12, expression(epsilon == -1), cex=.7, srt=-45)
text(1.2e+02, 1e+12, expression(epsilon == -2), cex=.7, srt=-45)
text(0.2e+02, .25e+12, expression(epsilon == -3), cex=.7, srt=-45)
text(0.15e+02, 1.3e+10, expression(epsilon == -4), cex=.7, srt=-45)

text(0.15e+05, .5e+12, "20 MWatt", cex=.7, srt=45)
text(0.15e+05, .5e+10, "0.2 MWatt", cex=.7, srt=45)
#text(0.15e+05, .5e+08, "0.02 MWatt", cex=.7, srt=45)


## GENERATE CONSTANT POWER AND ISO-EXASCALAR LINES FOR THE GRAPH

## These are tbe hashed lines on the graph representing "constant power" and "constant exascalar"
## The approach here is to define functions that create two endpoints for the line segments that 
##    depend on power and exascalar, respectively

isopowerline <- function(megawatts, efficiencyrangelow=10*.9, efficiencyrangehigh=100000*1.1, perfrangehigh = 2*10^12*1.1, perfrangelow = 0.9**5*10^6)
{matrix(c(efficiencyrangelow, efficiencyrangehigh, efficiencyrangelow*megawatts*10^6, efficiencyrangehigh*megawatts*10^6), ncol=2)
}

## Compute Graphical Lines for iso power
## the argument passed in isopowerline is in megawatts

for (i in -2:4) {
        lines(isopowerline(2*10^-i)[,1],isopowerline(2*10^-i)[,2], lwd=.5, lty=2)}

## add graphical lines for iso "exscalar"


ExaPerf <- 10^12           ##in Megaflops
ExaEff <- 10^12/(20*10^6)  ##in Megaflops/Watt

Exascale = ExaPerf*ExaEff

isoexaline <- function(exascalar, efficiencyrangelow=0.5*20, efficiencyrangehigh=2*50000, perfrangehigh = 2*10^12, perfrangelow = 0.5*5*10^7){
        matrix(c(Exascale/(10^(-exascalar*sqrt(2)))/perfrangehigh, efficiencyrangehigh, perfrangehigh, Exascale/(10^(-exascalar*sqrt(2)))/efficiencyrangehigh), ncol=2)
}

## Compute Graphical Lines for iso exascalar
## the argument passed is - log10 of exascalar

for (i in 0:7) {lines(isoexaline(-i)[,1], isoexaline(-i)[,2], lwd=.5, lty=2)}



```


###November 2014 Accelerators versus CPU-only


```{r "Tech Trend String Search", echo=FALSE}

TechTrendSearchString = "nvidia"
## use grepl to search for substring
bbb1 <- grepl(TechTrendSearchString, Nov14$computer, ignore.case=TRUE)

TechTrendSearchString = "phi"
## use grepl to search for substring
bbb2 <- grepl(TechTrendSearchString, Nov14$computer, ignore.case=TRUE)



TechTrendSearchString = "FirePro"
## use grepl to search for substring
bbb3 <- grepl(TechTrendSearchString, Nov14$computer, ignore.case=TRUE)



TechTrendSearchString = "PEZY"
## use grepl to search for substring
bbb4 <- grepl(TechTrendSearchString, Nov14$computer, ignore.case=TRUE)

bbb<-bbb1|bbb2|bbb3|bbb4

## get the locations
ccc <- which(bbb ==TRUE)
## subset Big Exascalar
Nov14x <- Nov14[ccc,]
```

In this analysis systems with GPUs made by NVidia, PEZY, AMD FirePro and Intel Xeon Phi are highlighted. (Respectful of copyrights). 

In all there are `r dim(Nov14x)[1]` systems out of the Top/Green500 highlighted. (`r round(100*dim(Nov14x)[1]/500,1) `%).


```{r, echo=FALSE, fig.align='center', fig.height=7.25, fig.width=6}
## EXASCALAR PLOT OVERLAYING TWO LISTS

## plots "reference" list first, then "list of current interest" is overlayed

plot(Nov14$mflopswatt ,
     Nov14$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2014 Exascalar", 
     col = "red",
     cex=1.1,
     pch=20, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))

par(new=TRUE)

plot(Nov14x$mflopswatt ,
     Nov14x$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2014 Exascalar", 
     col = "blue",
     cex=1.1,
     pch=19, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))


##add text to plots  (Some are commented out to clean up appearance, but left in for possible later convenience)

legend(4E1,1E11, legend = c("CPU","CPU/GPU"), pch=c(20,19), col = c("red", "blue"), cex=1)

text(1.0E4,
     5.E7, "Nov 2014", cex=.7, col="black", pos=3)
text(0.8E4,
     3.3E7, "data from Green500 and Top500 Lists", cex=.7, col="black", pos=3)


#text(0.35e+05, 2e+12, expression(epsilon == 0), cex=.7, srt=-45)
text(0.35e+04, 1e+12, expression(epsilon == -1), cex=.7, srt=-45)
text(1.2e+02, 1e+12, expression(epsilon == -2), cex=.7, srt=-45)
text(0.2e+02, .25e+12, expression(epsilon == -3), cex=.7, srt=-45)
text(0.15e+02, 1.3e+10, expression(epsilon == -4), cex=.7, srt=-45)

text(0.15e+05, .5e+12, "20 MWatt", cex=.7, srt=45)
text(0.15e+05, .5e+10, "0.2 MWatt", cex=.7, srt=45)
#text(0.15e+05, .5e+08, "0.02 MWatt", cex=.7, srt=45)


## GENERATE CONSTANT POWER AND ISO-EXASCALAR LINES FOR THE GRAPH

## These are tbe hashed lines on the graph representing "constant power" and "constant exascalar"
## The approach here is to define functions that create two endpoints for the line segments that 
##    depend on power and exascalar, respectively

isopowerline <- function(megawatts, efficiencyrangelow=10*.9, efficiencyrangehigh=100000*1.1, perfrangehigh = 2*10^12*1.1, perfrangelow = 0.9**5*10^6)
{matrix(c(efficiencyrangelow, efficiencyrangehigh, efficiencyrangelow*megawatts*10^6, efficiencyrangehigh*megawatts*10^6), ncol=2)
}

## Compute Graphical Lines for iso power
## the argument passed in isopowerline is in megawatts

for (i in -2:4) {
        lines(isopowerline(2*10^-i)[,1],isopowerline(2*10^-i)[,2], lwd=.5, lty=2)}

## add graphical lines for iso "exscalar"


ExaPerf <- 10^12           ##in Megaflops
ExaEff <- 10^12/(20*10^6)  ##in Megaflops/Watt

Exascale = ExaPerf*ExaEff

isoexaline <- function(exascalar, efficiencyrangelow=0.5*20, efficiencyrangehigh=2*50000, perfrangehigh = 2*10^12, perfrangelow = 0.5*5*10^7){
        matrix(c(Exascale/(10^(-exascalar*sqrt(2)))/perfrangehigh, efficiencyrangehigh, perfrangehigh, Exascale/(10^(-exascalar*sqrt(2)))/efficiencyrangehigh), ncol=2)
}

## Compute Graphical Lines for iso exascalar
## the argument passed is - log10 of exascalar

for (i in 0:7) {lines(isoexaline(-i)[,1], isoexaline(-i)[,2], lwd=.5, lty=2)}



```




###November 2012 Accelerators versus CPU-only


For comparison, this is the same analysis for the 2012 Exascalar version of the Top/Green500. Note that significantly fewer GPU systems were in the population.  

```{r "Tech Trend String Search2", echo=FALSE}

TechTrendSearchString = "nvidia"
## use grepl to search for substring
bbb1 <- grepl(TechTrendSearchString, Nov12$computer, ignore.case=TRUE)

TechTrendSearchString = "phi"
## use grepl to search for substring
bbb2 <- grepl(TechTrendSearchString, Nov12$computer, ignore.case=TRUE)



TechTrendSearchString = "FirePro"
## use grepl to search for substring
bbb3 <- grepl(TechTrendSearchString, Nov12$computer, ignore.case=TRUE)



TechTrendSearchString = "PEZY"
## use grepl to search for substring
bbb4 <- grepl(TechTrendSearchString, Nov12$computer, ignore.case=TRUE)

bbb<-bbb1|bbb2|bbb3|bbb4

## get the locations
ccc <- which(bbb ==TRUE)
## subset Big Exascalar
Nov12x <- Nov12[ccc,]
```

Systems with GPUs made by NVidia, PEZY, AMD FirePro and Intel Xeon Phi are highlighted. (Respectful of copyrights). 

In all there are `r dim(Nov12x)[1]` systems out of the Top/Green500 highlighted. (`r round(100*dim(Nov12x)[1]/500,1) `%).

```{r, echo=FALSE, fig.align='center', fig.height=7.25, fig.width=6}
## EXASCALAR PLOT OVERLAYING TWO LISTS

## plots "reference" list first, then "list of current interest" is overlayed

plot(Nov12$mflopswatt ,
     Nov12$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2012 Exascalar", 
     col = "red",
     cex=1.1,
     pch=20, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))

par(new=TRUE)

plot(Nov12x$mflopswatt ,
     Nov12x$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "",
     ylab = "", 
     main = "", 
     col = "blue",
     cex=1.1,
     pch=19, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))


##add text to plots  (Some are commented out to clean up appearance, but left in for possible later convenience)

legend(4E1,1E11, legend = c("CPU","CPU/GPU"), pch=c(20,19), col = c("red", "blue"), cex=1)

text(1.0E4,
     5.E7, "Nov 2012", cex=.7, col="black", pos=3)
text(0.8E4,
     3.3E7, "data from Green500 and Top500 Lists", cex=.7, col="black", pos=3)


#text(0.35e+05, 2e+12, expression(epsilon == 0), cex=.7, srt=-45)
text(0.35e+04, 1e+12, expression(epsilon == -1), cex=.7, srt=-45)
text(1.2e+02, 1e+12, expression(epsilon == -2), cex=.7, srt=-45)
text(0.2e+02, .25e+12, expression(epsilon == -3), cex=.7, srt=-45)
text(0.15e+02, 1.3e+10, expression(epsilon == -4), cex=.7, srt=-45)

text(0.15e+05, .5e+12, "20 MWatt", cex=.7, srt=45)
text(0.15e+05, .5e+10, "0.2 MWatt", cex=.7, srt=45)
#text(0.15e+05, .5e+08, "0.02 MWatt", cex=.7, srt=45)


## GENERATE CONSTANT POWER AND ISO-EXASCALAR LINES FOR THE GRAPH

## These are tbe hashed lines on the graph representing "constant power" and "constant exascalar"
## The approach here is to define functions that create two endpoints for the line segments that 
##    depend on power and exascalar, respectively

isopowerline <- function(megawatts, efficiencyrangelow=10*.9, efficiencyrangehigh=100000*1.1, perfrangehigh = 2*10^12*1.1, perfrangelow = 0.9**5*10^6)
{matrix(c(efficiencyrangelow, efficiencyrangehigh, efficiencyrangelow*megawatts*10^6, efficiencyrangehigh*megawatts*10^6), ncol=2)
}

## Compute Graphical Lines for iso power
## the argument passed in isopowerline is in megawatts

for (i in -2:4) {
        lines(isopowerline(2*10^-i)[,1],isopowerline(2*10^-i)[,2], lwd=.5, lty=2)}

## add graphical lines for iso "exscalar"


ExaPerf <- 10^12           ##in Megaflops
ExaEff <- 10^12/(20*10^6)  ##in Megaflops/Watt

Exascale = ExaPerf*ExaEff

isoexaline <- function(exascalar, efficiencyrangelow=0.5*20, efficiencyrangehigh=2*50000, perfrangehigh = 2*10^12, perfrangelow = 0.5*5*10^7){
        matrix(c(Exascale/(10^(-exascalar*sqrt(2)))/perfrangehigh, efficiencyrangehigh, perfrangehigh, Exascale/(10^(-exascalar*sqrt(2)))/efficiencyrangehigh), ncol=2)
}

## Compute Graphical Lines for iso exascalar
## the argument passed is - log10 of exascalar

for (i in 0:7) {lines(isoexaline(-i)[,1], isoexaline(-i)[,2], lwd=.5, lty=2)}



```








###Intel Xeon 

We can also look at other architectural influencers. This analysis looks at which computers rely on Intel Xeon for the CPU. (copyrights respected)


```{r "Tech Trend String SearchXeon", echo=FALSE}

TechTrendSearchString = "xeon"
## use grepl to search for substring
bbb1 <- grepl(TechTrendSearchString, Nov14$computer, ignore.case=TRUE)

bbb<-bbb1

## get the locations
ccc <- which(bbb ==TRUE)
## subset Big Exascalar
Nov14x <- Nov14[ccc,]
```

In all there are `r dim(Nov14x)[1]` systems out of the Top/Green500 highlighted. (`r round(100*dim(Nov14x)[1]/500,1) `%).


```{r, echo=FALSE, fig.align='center', fig.height=7.25, fig.width=6}
## EXASCALAR PLOT OVERLAYING TWO LISTS

## plots "reference" list first, then "list of current interest" is overlayed

plot(Nov14$mflopswatt ,
     Nov14$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2014 Exascalar", 
     col = "red",
     cex=1.1,
     pch=20, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))

par(new=TRUE)

plot(Nov14x$mflopswatt ,
     Nov14x$rmax*10^3, 
     log="xy", 
     asp = 4/3.2, 
     xlab = "Efficiency (mflops/watt)",
     ylab = "Performance (mflops)", 
     main = "November 2014 Exascalar", 
     col = "blue",
     cex=1.1,
     pch=19, 
     xlim=c(10,100000), 
     ylim=c(5*10^7,2*10^12))


##add text to plots  (Some are commented out to clean up appearance, but left in for possible later convenience)

legend(4E1,1E11, legend = c("Other","Intel(R) Xeon(R)"), pch=c(20,19), col = c("red", "blue"), cex=1)

text(1.0E4,
     5.E7, "Nov 2014", cex=.7, col="black", pos=3)
text(0.8E4,
     3.3E7, "data from Green500 and Top500 Lists", cex=.7, col="black", pos=3)


#text(0.35e+05, 2e+12, expression(epsilon == 0), cex=.7, srt=-45)
text(0.35e+04, 1e+12, expression(epsilon == -1), cex=.7, srt=-45)
text(1.2e+02, 1e+12, expression(epsilon == -2), cex=.7, srt=-45)
text(0.2e+02, .25e+12, expression(epsilon == -3), cex=.7, srt=-45)
text(0.15e+02, 1.3e+10, expression(epsilon == -4), cex=.7, srt=-45)

text(0.15e+05, .5e+12, "20 MWatt", cex=.7, srt=45)
text(0.15e+05, .5e+10, "0.2 MWatt", cex=.7, srt=45)
#text(0.15e+05, .5e+08, "0.02 MWatt", cex=.7, srt=45)


## GENERATE CONSTANT POWER AND ISO-EXASCALAR LINES FOR THE GRAPH

## These are tbe hashed lines on the graph representing "constant power" and "constant exascalar"
## The approach here is to define functions that create two endpoints for the line segments that 
##    depend on power and exascalar, respectively

isopowerline <- function(megawatts, efficiencyrangelow=10*.9, efficiencyrangehigh=100000*1.1, perfrangehigh = 2*10^12*1.1, perfrangelow = 0.9**5*10^6)
{matrix(c(efficiencyrangelow, efficiencyrangehigh, efficiencyrangelow*megawatts*10^6, efficiencyrangehigh*megawatts*10^6), ncol=2)
}

## Compute Graphical Lines for iso power
## the argument passed in isopowerline is in megawatts

for (i in -2:4) {
        lines(isopowerline(2*10^-i)[,1],isopowerline(2*10^-i)[,2], lwd=.5, lty=2)}

## add graphical lines for iso "exscalar"


ExaPerf <- 10^12           ##in Megaflops
ExaEff <- 10^12/(20*10^6)  ##in Megaflops/Watt

Exascale = ExaPerf*ExaEff

isoexaline <- function(exascalar, efficiencyrangelow=0.5*20, efficiencyrangehigh=2*50000, perfrangehigh = 2*10^12, perfrangelow = 0.5*5*10^7){
        matrix(c(Exascale/(10^(-exascalar*sqrt(2)))/perfrangehigh, efficiencyrangehigh, perfrangehigh, Exascale/(10^(-exascalar*sqrt(2)))/efficiencyrangehigh), ncol=2)
}

## Compute Graphical Lines for iso exascalar
## the argument passed is - log10 of exascalar

for (i in 0:7) {lines(isoexaline(-i)[,1], isoexaline(-i)[,2], lwd=.5, lty=2)}

```


